Repeatability of Pre-Flashover Fire Patterns on Gypsum Wallboard

### 2016-09-17 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
Ch X

As you begin to examine the scene, you walk around the exterior and note that on the front of the end unit, townhouse, there is fire damage on the exterior above the front door opening that extends across the ceiling of the porch.  The is some thermal damage and smoke deposits on the from of the siding on the porch.  There outer pane of the living room window is for the most part still intact. you notice that three of the four windows on the second floor have been removed.  On the ``D side'' of the townhome there are no openings and no smoke or fire damage is present.  On the rear or ``Side C'' of the house the door and windows are intact but open.  There is no smoke stains or thermal damage around any of the openings on the rear. No sign of smoke alarms in the home.

You take some notes and photograph the exterior of the structure, the fire department is not repacking their equipment and are preparing to leave when you enter through the front doorway.  The living room has extensive fire damage.  Looking to your right, two chairs infront of the living room window have some thermal damage.  There are a clear horizontal line of demarcation the large wood TV stand in the corner.  As your eyes move from the corner back toward the rear of the townhome and the kitchen, you notice that the elevation of the line of demarcation is increasing.  This may be an indicator that the origin of the fire is closer to the front of the living room.  Between the two wood cabinets on the ``D wall'' of the living room was a wood framed sofa.  The cushions have thermal damage, from pyrolysis.  It is not clear that they were on fire as there is not thermal damage on the wall directly above the top of the sofa cushions that would indicate the presence of a thermal plume.  Still standing in the entry doorway, you turn your attention to the remains of an upholstered sofa and chair.  The wall behind the sofa and chair has extensive thermal damage.  The wall extends from the opening to the dining/kitchen area of the house to the open stairway that goes to the upper level.  You move to the dining area and kitchen area.  There is plenty of smoke damage high in the space, but there are no indications of fire damage due to combustion.  You return to the living room and walk up the stairs.  Their is evidence of paint and paper burned off of the gypsum wallboard in the stairway.  The fire damage doe not extend beyond the top of the stairs.  there is smoke damage in the upstairs hallway an in the front bedrooms. The further you move from the stairs the less damage you find.

You head back down the stairs to further examine the living room which appears to be the room of fire origin. Your are focusing on the area with the most damage, the area with the sofa and chair.  As you are looking toward the ``B wall'' of the living room you note that the right side of the chair is badly burned and it is burned from the floor level up.  The same is true from the left side of the sofa.  The corner formed by the wall behind and the wall of the stairway has a ``V pattern'' burned into it, from the floor and extending up toward the ceiling.  As you look at the debris surrounding the chair, you notice the remains of an electric heater.  The wire from the heater extends toward the wall outlet behind the sofa.

You have no developed your first hypothesis about the origin of the fire.  A radiant style, electric heater was turned on and left until it ignited the fabric on the sofa or chair.  You collect the remains of the heater for further examination to have a lab determine of the heater switch was in the ``on'' position.  No you need to start testing your hypothesis.  Is everything consistent with the other fire damage?

There are different damage patterns behind the sofa.  On the right side there is fire damage on the wall behind the right side back cushion of the sofa. The damaged area is distinct from the other fire damage that exists on the wall behind the sofa. There are pieces of charred newspaper near the right arm of the sofa.  The burn mark on the top of the wood coffee looks like a arrow that is pointing toward the corner.  The char on the legs of the coffee table is very directional and consistent with the heat coming from the corner.  Looking at the inside area of the sofa, you see that there is charred wood on the inside surface of the sofa frame while the outer portions of the wood frame are not charred.  This would indicate that the fire was inside the sofa at some point.  Further is appears that all of the foam cushioning is burnt off of the right side of the sofa, while there is a piece of polyurethane foam cushion on the left side of the sofa.  However that foam could have gotten moved during the fire suppression.

@@

For simulations involving buoyant plumes, a measure of how well the flow field is resolved is given by the non-dimensional expression $D^*/\dx$, where $D^*$ is a characteristic
fire diameter

\begin{equation}
 D^* = \left(
     \frac{\dQ}{\rho_\infty \, c_p \, T_\infty \, \sqrt{g} }
     \right)^\frac{2}{5} 
 \end{equation}

and $\dx$ is the nominal size of a mesh cell.

The characteristic fire diameter is related to the characteristic fire size via the
relation $Q^* = (D^*/D)^{5/2}$, where $D$ is the physical diameter of the
fire.}. The quantity, $\dQ$, is the total heat release rate of the fire. If it changes over time, you should consider the corresponding change in resolution. The quantity $D^*/\dx$ can be thought of as the number of computational cells spanning the characteristic (not necessarily the physical) diameter of the fire.

@@

@@

@@

>>>>

As you begin to examine the scene, you walk around the exterior and note that on the front of the end unit townhouse, there is fire damage on the exterior above the front door opening that extends across the ceiling of the porch.  The is some thermal damage and smoke deposits on the from of the siding on the porch.  The outer pane of the living room window is for the most part still intact. you notice that three of the four windows on the second floor have been removed.  On ``side D'' of the townhome there are no openings and no smoke or fire damage is present.  On the rear or ``side C'' of the house the door and windows are intact but open.  There is no smoke stains or thermal damage around any of the openings on the rear. No sign of smoke alarms in the home.

You take some notes and photograph the exterior of the structure, the fire department is not repacking their equipment and are preparing to leave when you enter through the front doorway.  The living room has extensive fire damage.  Looking to your right, two chairs in front of the living room window have some thermal damage.  There are a clear horizontal line of demarcation the large wood TV stand in the corner.  As your eyes move from the corner back toward the rear of the townhome and the kitchen, you notice that the elevation of the line of demarcation is increasing.  This may be an indicator that the origin of the fire is closer to the front of the living room.  Between the two wood cabinets on the ``D wall'' of the living room was a wood framed sofa.  The cushions have thermal damage, from pyrolysis.  It is not clear that they were on fire as there is not thermal damage on the wall directly above the top of the sofa cushions that would indicate the presence of a thermal plume.  Still standing in the entry doorway, you turn your attention to the remains of an upholstered sofa and chair.  The wall behind the sofa and chair has extensive thermal damage.  The wall extends from the opening to the dining/kitchen area of the house to the open stairway that goes to the upper level.  You move to the dining area and kitchen area.  There is plenty of smoke damage high in the space, but there are no indications of fire damage due to combustion.  You return to the living room and walk up the stairs.  There is evidence of paint and paper burned off of the gypsum wallboard in the stairway.  The thermal damage does not extend beyond the top of the stairs.  There is smoke damage in the upstairs hallway and in the front bedrooms. The further you move from the stairs the less damage you find.

You head back down the stairs to further examine the living room which appears to be the room of fire origin. Your are focusing on the area with the most damage, the area with the sofa and chair.  As you are looking toward the ``B wall'' of the living room you note that the right side of the chair is badly burned and it is burned from the floor level up.  The same is true from the left side of the sofa.  The corner formed by the wall behind the chair and the wall of the stairway has a ``V pattern'' burned into it, from the floor and extending up toward the ceiling.  As you look at the debris surrounding the chair, you notice the remains of an electric heater.  The burned wire from the heater extends toward the wall outlet behind the sofa.

You have now developed your first hypothesis about the origin of the fire.  A radiant style, electric heater was turned on and left until it ignited the fabric on the sofa or chair.  You collect the remains of the heater for further examination to have a lab determine if the heater switch was in the ``on'' position.  Now you need to start testing your hypothesis.  Is everything consistent with the other fire damage?

There are different damage patterns behind the sofa.  On the right side there is fire damage on the wall behind the right side back cushion of the sofa. The damaged area is distinct from the other fire damage that exists on the wall behind the sofa. There are pieces of charred newspaper near the right arm of the sofa.  The burn mark on the top of the wood coffee table looks like a arrow that is pointing toward the corner.  The char on the legs of the coffee table is very directional and consistent with the heat coming from the corner.  Looking at the inside area of the sofa, you see that there is charred wood on the inside surface of the sofa frame while the outer portions of the wood frame are not charred.  This would indicate that the fire was inside the sofa at some point.  Further is appears that all of the foam cushioning is burnt off of the right side of the sofa, while there is a piece of polyurethane foam cushion on the left side of the sofa.  However that foam could have gotten moved during the fire suppression.

@@

@@

While you are reviewing the scene you meet up with the fire captain and begin asking him about the fire.  The captain tells you that the doors and windows were closed upon arrival.  When the front door was opened heavy smoke

@@

degradation of gypsum board full scale

Use of black paint

Calcination depth

Larger variation of heat release rates

Realistic fuels

@@

Repeatibilty
Energy to Damage
Flame Height
Modeling
